Log or not

### 2021-10-23 analysis/Analysis.Rmd
add plots and clean analysis script

n <- 100
quantiles <- c(0.01, 0.025, seq(0.05, 0.95, by = 0.05), 0.975, 0.99)

@@

scenarios <- scenarios |>
  inner_join(grids) |>
  mutate(name = paste0(error_type, "_error_", error_var,  
                       ifelse(equal_mean, "_equal_mean_", "_unequal_mean_"), 
                       ifelse(equal_var, "equal_var", "unequal_var")))

@@

```{r}
# helper functions
# ----------------

@@

# helper function to plot scores
plot_scores <- function(scores) {
  scores |>
    mutate("True parameters" = 
             paste0("mean: ", mu_true, ", sd: ", sigma_true)) |>
    ggplot(aes(x = error, y = interval_score, 
               colour = `True parameters`, group = `True parameters`)) + 
    geom_line() + 
    scale_color_brewer(palette = "Set1", name = "True parameters: ") + 
    facet_wrap(~ scale, scale = "free") + 
    theme_minimal() + 
    theme(legend.position = "bottom") 
}

@@

```{r}
i = 1
equal_mean <- scenarios[i, ]$truth_mean_equal
equal_var <- scenarios[i, ]$truth_var_equal
grid <- unlist(scenarios[i, ]$grid)
error_var <- scenarios[i, ]$error_var
error_type <- scenarios[i, ]$error_type
scenario_name <- scenarios[i, ]$name

sigma_true <- c(sigma, sigma + (as.numeric(!equal_var) * 10))
mu_true <- c(mu, mu + as.numeric(!equal_mean) * 1000)

@@

title <- paste(error_type, "error of", error_var, "for two variables with", 
               ifelse(equal_mean, "equal", "unequal"), 
               "mean and", 
               ifelse(equal_var, "equal", "unequal"), "variance") |>
    str_to_sentence()

xlab <- paste(error_type, "error on", error_var) |> str_to_sentence()

plot_scores(combined_scores) + 
  ggtitle(title) + 
  labs(y = "Interval Score", 
       x = xlab)

ggsave(here("output", "figures", paste0(scenario_name, ".png")))
```

Simulate data from two normal distributions with $\mu_{1,2} = 1000$ and $\sigma_1 = 0.1$ and $\sigma_2 = 1$.

```{r}
sigma <- c(1, 100)
mu_true <- c(1000, 2000)

true_values_1 <- 
  data.frame(
    true_value = rnorm(n = n, mean = mu_true[1], sd = sigma[1]),
    n = 1:n,
    sigma_true = sigma[1]
  )

true_values_2 <- 
  data.frame(
    true_value = rnorm(n = n, mean = mu_true[1], sd = sigma[2]),
    n = 1:n,
    sigma_true = sigma[2]
  )

@@

Define different predictive distributions with mean 1000 and standard deviations between 0 and 10.

```{r}
mu_hat <- 1000

forecasts_1 <- create_forecast(grid_sigma_mult, sigma_true = sigma[1]) 
forecasts_2 <- create_forecast(grid_sigma_mult, sigma_true = sigma[2])

forecasts_mu_1 <- create_forecast(grid_mu_mult, sigma_true = sigma[1], 
                                 mu_true = 1000, var = "mu") 
forecasts_mu_2 <- create_forecast(grid_mu_mult, sigma_true = sigma[2], 
                                 mu_true = 1000, var = "mu") 
```

Join these and create a version with predictions and true values logged. 
```{r}
data_1 <- inner_join(forecasts_1, true_values_1)
data_2 <- inner_join(forecasts_2, true_values_2)

data_mu_1 <- inner_join(forecasts_mu_1, true_values_1)
data_mu_2 <- inner_join(forecasts_mu_2, true_values_2)

log_data_1 <- create_log_data(data_1)
log_data_2 <- create_log_data(data_2)

log_data_mu_1 <- create_log_data(data_mu_1)
log_data_mu_2 <- create_log_data(data_mu_2)
```

Score forecasts.

```{r}
scores_1 <- score_forecasts(data_1)
scores_2 <- score_forecasts(data_2)

log_scores_1 <- score_forecasts(log_data_1)
log_scores_2 <- score_forecasts(log_data_2)

scores_mu_1 <- score_forecasts(data_mu_1)
scores_mu_2 <- score_forecasts(data_mu_2)

log_scores_mu_1 <- score_forecasts(log_data_mu_1)
log_scores_mu_2 <- score_forecasts(log_data_mu_2)

joined_scores_mu <- join_scores(scores_mu_1, scores_mu_2)
joined_log_scores_mu <- join_scores(log_scores_mu_1, log_scores_mu_2, scale = "log")
combined_scores_mu <- rbind(joined_scores_mu, joined_log_scores_mu)

@@

Create a plot for data on log scale and natural scale.

```{r}

## Scenario: Multiplicative error on the standard deviation - identical means

```{r}
plot_scores(combined_scores) +
    labs(y = "WIS", x = "sd of predictive distribution / sd of data distribution") 
```

*Plot shows forecasts where the mean of the forecast distribution is equal to the mean of the data-generating distribution. The variance of the forecast distribution is multiplied by an error factor between 0.5 and 10.*

Same plot with y axis on a log scale
```{r}
plot_scores(combined_scores) +
    labs(y = "WIS", x = "sd of predictive distribution / sd of data distribution") + 
  ggplot2::scale_y_continuous(trans = "log10")
```

What we learn from these plots:

- A perfect forecast that gets the mean as well as the variance right will incur higher scores if the true variance is higher. 
- getting the variance wrong (in multiplicative terms) is equally costly regardless of what the true variance is (curves look identical on a log scale plot)
- Those two effects are exactly the same on the linear as on the log scale? Not sure I understand this

## Multiplicative error on the mean - identical means

```{r}
plot_scores(combined_scores_mu) +
    labs(y = "WIS", x = "mu of predictive distribution / mu of data distribution")
```

*Plot shows forecasts where the standard deviation of the forecast distribution is equal to the standard deviation of the data-generating distribution. The mean of the forecast distribution is multiplied by an error factor between 0.5 and 10.*

@@

```{r}
combined_scores_mu |>
  group_by(error, scale) |>
  summarise(mean_score = mean(interval_score))
```

>>>>

n <- 100
quantiles <- c(0.01, 0.025, seq(0.05, 0.95, by = 0.05), 0.975, 0.99)
rerun_simulation <- FALSE

@@

scenarios <- scenarios |>
  inner_join(grids) |>
  mutate(name = paste0(error_type, "_error_", error_var,  
                       ifelse(truth_mean_equal, "_equal_mean_", "_unequal_mean_"), 
                       ifelse(truth_var_equal, "equal_var", "unequal_var")), 
         title = paste(error_type, "error of", error_var, "for two variables with", 
                       ifelse(truth_mean_equal, "equal", "unequal"), 
                       "mean and", 
                       ifelse(truth_var_equal, "equal", "unequal"), "variance") |>
           str_to_sentence())

scenarios |>
  select(truth_mean_equal, truth_var_equal, error_var, error_type)

@@

```{r, eval = rerun_simulation}
# helper functions
# ----------------

@@

# helper function to wrap a long title
wrap_title <- function(x, ...) {
  paste(strwrap(x, ...), collapse = "\n")
}

# helper function to plot scores
plot_scores <- function(scores) {
  scores |>
    mutate("True parameters" = 
             paste0("mean: ", mu_true, ", sd: ", sigma_true)) |>
    group_by(`True parameters`, error, scale) |>
    summarise(interval_score = mean(interval_score)) |>
    ggplot(aes(x = error, y = interval_score, 
               colour = `True parameters`, group = `True parameters`)) + 
    geom_line() + 
    scale_color_brewer(palette = "Set1", name = "True parameters: ") + 
    facet_wrap(~ scale, scale = "free") + 
    theme_minimal() + 
    theme(legend.position = "bottom") 
}

@@

```{r eval = rerun_simulation}
for (i in 1:nrow(scenarios)) {
  equal_mean <- scenarios[i, ]$truth_mean_equal
  equal_var <- scenarios[i, ]$truth_var_equal
  grid <- unlist(scenarios[i, ]$grid)
  error_var <- scenarios[i, ]$error_var
  error_type <- scenarios[i, ]$error_type
  scenario_name <- scenarios[i, ]$name
  scenario_title <- scenarios[i, ]$title

  sigma_true <- c(sigma, sigma + (as.numeric(!equal_var) * 90))
  mu_true <- c(mu, mu + as.numeric(!equal_mean) * 1000)

@@

  plot_scores(combined_scores) + 
    ggtitle(wrap_title(scenario_title, width = 60)) + 
    labs(y = "Interval Score", 
         x = paste(error_type, "error on", error_var) |> 
           str_to_sentence())

  ggsave(here("output", "figures", paste0(scenario_name, ".png")))
}

@@

---

```{r}
i = 1
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 2
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 3
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 4
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 5
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 6
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 7
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 8
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 9
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))

@@

@@

---

```{r}
i = 10
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 11
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

- A perfect forecast that gets the mean as well as the variance right will incur higher scores if the true variance is higher. 
- getting the variance wrong (in multiplicative terms) is equally costly regardless of what the true variance is (curves look identical on a log scale plot, I checked that previously)
- Those two effects are exactly the same on the linear as on the log scale? Not sure I understand this

---

```{r}
i = 12
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 13
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 14
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 15
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 16
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```
